Eco-evolutionary dynamics of finite populations from first principles

### 2023-02-18 ind_files/99_Appendix_C.tex
Correct minor calculation error

&+ 2\textrm{Cov}\left(\frac{\partial f}{\partial t},f\right)dt + M_{\sigma^2_f}(\mathbf{p},N_K)dt + dB_{\sigma^2_{f}}

@@

&- \frac{1}{KN_K}\left[ \ \textrm{Cov}\left(\tau,(f - \overline{f})^2\right) - \textrm{Cov}\left(\tau,f^2 - {\left(\overline{f}\right)}^2\right) \ \right]dt\\[12pt]
& + 2\textrm{Cov}\left(\frac{\partial f}{\partial t},f\right)dt + M_{\sigma^2_f}(\mathbf{p},N_K)dt + dB_{\sigma^2_{f}}

@@

>>>>

&+ 2\textrm{Cov}\left(f, \frac{\partial f}{\partial t}\right)dt + M_{\sigma^2_f}(\mathbf{p},N_K)dt + dB_{\sigma^2_{f}}

@@

&- \frac{1}{KN_K}\left[ \ \textrm{Cov}\left(\tau,(f - \overline{f})^2\right) + \textrm{Cov}\left(\tau,f^2 - {\left(\overline{f}\right)}^2\right) \ \right]dt\\[12pt]
& + 2\textrm{Cov}\left(f, \frac{\partial f}{\partial t}\right)dt + M_{\sigma^2_f}(\mathbf{p},N_K)dt + dB_{\sigma^2_{f}}

@@

To simplify the covariance terms of the first line of the RHS, we observe that
\textrm{Cov}\left(w, {\left(\overline{f}\right)}^2\right) &= \overline{\left(w{\left(\overline{f}\right)}^2\right)} - \overline{w}\overline{\left({\left(\overline{f}\right)}^2\right)}\nonumber\\
&= {\left(\overline{f}\right)}^2\sum\limits_{i=1}^{m}w_ip_i - \overline{w}{\left(\overline{f}\right)}^2\sum\limits_{i=1}^{m}p_i\nonumber\\
&= {\left(\overline{f}\right)}^2\overline{w} - \overline{w}{\left(\overline{f}\right)}^2 = 0\label{cov_term_1_for_variances}
and
\textrm{Cov}\left(\tau,f^2 - {\left(\overline{f}\right)}^2\right) &= \overline{\tau\left(f^2 - {\left(\overline{f}\right)}^2\right)} - \overline{\tau}\left(\overline{f^2 - {\left(\overline{f}\right)}^2}\right)\nonumber\\
&= \overline{\tau f^2} - \overline{\tau}{\left(\overline{f}\right)}^2  - \overline{\tau}\overline{f^2} + \overline{\tau}{\left(\overline{f}\right)}^2\nonumber\\
&=\overline{\tau f^2} - \overline{\tau}\overline{f^2} = \textrm{Cov}(\tau,f^2)\label{cov_term_2_for_variances}
Substituting \eqref{cov_term_1_for_variances} and \eqref{cov_term_2_for_variances} into \eqref{intermediate_2_for_variances}, we thus arrive at:
\label{intermediate_3_for_variances}
d\sigma^2_{f} &= \textrm{Cov}\left(w,(f - \overline{f})^2\right)dt\\
&- \frac{1}{KN_K}\left[ \ \textrm{Cov}\left(\tau,(f - \overline{f})^2\right) + \textrm{Cov}(\tau,f^2) \ \right]dt\\
& + 2\textrm{Cov}\left(f, \frac{\partial f}{\partial t}\right)dt + M_{\sigma^2_f}(\mathbf{p},N_K)dt + dB_{\sigma^2_{f}}
Covariance with $(f-\overline{f})^2$ is a measure of covariance with the existing `spread' of $f_i$ around the population mean value $\overline{f}$ and is thus readily understood to represent existing variation. However, the $\textrm{Cov}(\tau,f^2)$ term is difficult to interpret since it depends on the square of $f$, which may not have any direct biological interpretation. We will convert this into a form whose interpretation is more clear
% by repeatedly adding and subtracting a quantity from our expression (these quantities are shown in red below for clarity), noting that this leaves the net value unchanged. We have:
% \begin{align}
% \textrm{Cov}(\tau, f^2) &= \overline{\tau f^2} - \overline{\tau}\overline{f^2}\nonumber\\
% &= \overline{\tau f^2} {\color{red} + \overline{\tau{\left(\overline{f}\right)}^2} - \overline{\tau f \overline{f}}}  - \overline{\tau}\overline{f^2} \nonumber\\
% &= \overline{\tau f^2} + \overline{\tau{\left(\overline{f}\right)}^2} - 2\overline{\tau f \overline{f}}\nonumber\\
% &= \overline{\tau\left(f-\overline{f}\right)^2}\nonumber\\
% &= \overline{\tau\left(f-\overline{f}\right)^2} {\color{red}- \overline{\tau}{\left(\overline{f}\right)}^2 + 2 \overline{\tau}\overline{f\overline{f}} - \overline{\tau}{\left(\overline{f}\right)}^2}\nonumber\\
% &= \overline{\tau\left(f-\overline{f}\right)^2} {\color{red} - \overline{\tau}\overline{f^2}} -\overline{\tau}\overline{{\left(\overline{f}\right)}^2} + 2 \overline{\tau}\overline{f\overline{f}} {\color{red} + \overline{\tau}\overline{f^2}} - \overline{\tau}{\left(\overline{f}\right)}^2\nonumber\\
% &=  \overline{\tau\left(f-\overline{f}\right)^2} - \overline{\tau}\overline{\left(f-\overline{f}\right)^2} + \overline{\tau}\left(\overline{f^2} - {\left(\overline{f}\right)}^2\right)\nonumber\\
% &= \textrm{Cov}(\tau, \left(f-\overline{f}\right)^2) + \overline{\tau}\sigma^2_{f}\label{cov_to_var_sub_for_variances}
% \end{align}
% Finally, substituting \eqref{cov_to_var_sub_for_variances} into \eqref{intermediate_3_for_variances}, we see that the rate of change of the variance of any type-level quantity $f$ in the population satisfies:
% \begin{equation}
% \label{stochastic_Price_variance}
% \begin{aligned}
% d\sigma^2_{f} &= \textrm{Cov}\left(w,(f - \overline{f})^2\right)dt - \frac{1}{KN_K}\left[ \ 2\textrm{Cov}\left(\tau,(f - \overline{f})^2\right) + \overline{\tau}\sigma^2_{f}(t) \ \right]dt\\[12pt]
% & + 2\textrm{Cov}\left(\frac{\partial f}{\partial t},f\right)dt + M_{\sigma^2_f}(\mathbf{p},N_K)dt + dB_{\sigma^2_{f}}
% \end{aligned}
% \end{equation}
% In the case of one-dimensional quantitative traits, an infinite-dimensional version of \eqref{stochastic_Price_variance} has been rigorously derived \citep{week_white_2021} in a recent paper for the case where $f$ is the phenotypic value of the quantitative trait - Equation (21b) of \cite{week_white_2021} is precisely the $m \to \infty$ version of my equation \eqref{stochastic_Price_variance} when $f$ is the phenotypic value of the quantitative trait under consideration.
by using the fact that $f^2 = (f-\overline{f})^2 - \overline{f}^2 + 2f\overline{f}$. we have
\textrm{Cov}(\tau, f^2) &= \textrm{Cov}(\tau, (f-\overline{f})^2) + \textrm{Cov}(\tau, - {\left(\overline{f}\right)}^2) + 2f\overline{f}\nonumber\\
&=  \textrm{Cov}(\tau, (f-\overline{f})^2) \ \  - \underbrace{\textrm{Cov}(\tau, {\left(\overline{f}\right)}^2)}_{\substack{ =0 \\ {\text{by same argument as in \eqref{cov_term_1_for_variances}}}}} + \ \  2\overline{f}\textrm{Cov}(\tau, f)\label{cov_term_3_for_variances}
and thus, substituting \eqref{cov_term_3_for_variances} into \eqref{intermediate_3_for_variances},  we see that the rate of change of the variance of any type-level quantity $f$ in the population satisfies:
\label{stochastic_Price_variance}
d\sigma^2_{f} &= \textrm{Cov}\left(w,(f - \overline{f})^2\right)dt - \frac{2}{KN_K}\left[ \ \overline{f}\textrm{Cov}(\tau, f) +  \textrm{Cov}\left(\tau,(f - \overline{f})^2\right) \ \right]dt\\[12pt]
& + 2\textrm{Cov}\left(f,\frac{\partial f}{\partial t}\right)dt + M_{\sigma^2_f}(\mathbf{p},N_K)dt + dB_{\sigma^2_{f}}
\end{equation}.
